Some examples in extended logic counting classes

### 2016-12-26 LICS/appendix.tex
26 dic

First, for a given signature $\R$ and logic fragment $\LL$ we define a {\em valid form in $\LL$} (or just {\em valid form}). A formula $\alpha$ is in valid form in $\LL$ if (1) it is equal to $\varphi$, $1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{x}}\sa{\bar{X}}\varphi$ or $\sa{\bar{x}}\sa{\bar{X}}1$ for some $\LL$ formula $\varphi$ and some tuple of first-order (second-order) variables $\bar{x}$ ($\bar{X}$), or (2) if it is equal to $(\beta\add\gamma)$ where $\beta$ and $\gamma$ are formulas in valid form. Given a $\eqso(\loge{0})$ formula $\alpha$ over a signature $\R$ we will recursively construct a formula $\tau(\alpha)$ in valid form such that $\alpha(\enc(\A)) = \tau(\alpha)(\enc(\A))$ for each $\A\in\ostr[\R]$.

First, each natural constant $s$ in $\alpha$ is replaced by $(1\add 1\add\cdots\add 1)$ ($s$ times). If $\alpha = 1$ or $\alpha = \varphi \in \loge{0}$, then $\tau(\alpha) = \alpha$. We assume that every sub-formula of $\alpha$ is in valid form. If $\alpha = (\beta \add \gamma)$ then $\tau(\alpha) = \alpha$. If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$ this case separates in more cases: If $\alpha = \sa{x}\beta$ and $\beta$ is in one of the valid forms in (1), then $\tau(\alpha) = \alpha$. If $\alpha = \sa{X}\beta$ and $\beta$ is in the valid form $\sa{\bar{x}}\gamma$ then $\tau(\alpha) = \sa{\bar{x}}\sa{X}\gamma$; if $\beta$ it is in one of the remaining valid forms in (1), then $\tau(\alpha) = \alpha$. If $\alpha = \sa{x}\beta$ ($\alpha = \sa{X}\beta$) and $\beta = (\gamma_1\add\gamma_2)$, then $\tau(\alpha) = \tau(\sa{x}\gamma_1) \add \tau(\sa{x}\gamma_2)$ ($\tau(\alpha) = \tau(\sa{X}\gamma_1) \add \tau(\sa{X}\gamma_2)$). This covers all possible cases for $\alpha$. Furthermore, suppose that $\LL$ is closed under conjunction and suppose $\alpha = (\beta\mult\gamma)$. If $\beta = (\delta_1\add\delta_2)$ (wlog), then $\tau(\alpha) = \tau(\delta_1\mult\gamma)\add\tau(\delta_2\mult\gamma)$. If $\beta = \sa{\bar{x}}\delta$, then $\tau(\alpha) = \tau(\sa{\bar{x}}\tau(\delta\mult\gamma))$. If $\beta = 1$, then $\tau(\alpha) = \gamma$. If $\beta = \varphi$ and $\gamma = \psi$, then $\tau(\alpha) = \varphi\wedge\psi$. Let $\alpha$ be a $\eqso(\loge{0})$ formula over a signature $\R$. Consider the following procedure that receives $\enc(\A)$ as input, where $\A\in\ostr[\R]$ with domain $A$. Let $\tau(\alpha) = \beta_1\add\cdots\add\beta_m$. For each $\beta_i$, if $\beta_i = 1$ or $\beta_i = \varphi$, then $\sem{\beta_i}(\A)$ can be computed in polynomial time. If $\beta_i = \sa{\bar{x}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}$, if $\beta_i = \sa{\bar{X}}1$ then $\sem{\beta_i}(\A) = 2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, if $\beta_i = \sa{\bar{x}}\sa{\bar{X}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, all of which can be computed in polynomial time. Otherwise, $\beta_i$ is $\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, and then $\sem{\beta_i}(\A) = f_{\varphi}(\enc(\A))$, and $f_{\varphi}\in \#\loge{0}$ which implies that $f_{\varphi}(\enc(\A))$ can be computed in polynomial time \cite{DBLP:journals/jcss/SalujaST95}.

@@

Let $\alpha$ be a formula in $\eqso(\loge{1})$ over a signature $\R$. Let $\tau(\alpha) = \beta_1\add\beta_2\add\cdots\add\beta_m$ defined as in the proof of Theorem \ref{eqso-sigma-zero-in-fp}. This is, each $\beta_i$ is of the form $1$, $\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{x}}\sa{\bar{X}}1$ or $\sa{\bar{x}}\sa{\bar{X}}\varphi$, where $\varphi$ is a formula in $\loge{1}$ and $\bar{x}$ ($\bar{X}$) is a tuple of first-order (second-order) variables. For each $\beta_i$ we will construct a $\gamma_i$ such that $\gamma_i \in \eqso(\logu{1})$ and $\sem{\beta_i}(\A) = \sem{\gamma_i}(\A)$ for each $\A\in\ostr[\R]$. Since the forms $1$, $\sa{\bar{x}}1$, $\sa{\bar{X}}1$ and $\sa{\bar{x}}\sa{\bar{X}}1$ are already valid in $\eqso(\logu{1})$, we suppose $\beta_i$ is in one of the remaining forms. Without loss of generality, let $\beta_i = \sa{\bar{x}}\sa{\bar{X}}\exists\bar{y}\,\varphi(\bar{x},\bar{y},\bar{X})$ where $\varphi(\bar{x},\bar{y},\bar{X})$ is a quantifier-free formula. Similarly as in the proof by Saluja et al. \cite{DBLP:journals/jcss/SalujaST95}, we define $\gamma_i = \sa{\bar{x}}\sa{\bar{y}}\sa{\bar{X}}\varphi(\bar{x},\bar{y},\bar{X}) \wedge \forall\bar{z}(\varphi(\bar{x},\bar{z},\bar{X})\to(\bar{z} < \bar{y} \vee \bar{z} = \bar{y}))$ that fulfils the desired condition. The theorem holds true by using $\beta = \gamma_1\add\cdots\add\gamma_m$ which is in $\eqso(\logu{1})$

\subsection{Proof of Theorem \ref{one-sigma-zero}}

Towards a contradiction, suppose that there exists a constant function $f(\A) = s > 0$ which is expressible in $\E{0}$. This is, there is a formula $\psi(x,X)$ such that $f_{\psi(x,X)}(\A) = \vert \{ \langle e,P\rangle \mid \A\models\psi(e,P) \} \vert = s$ for each $\A\in\ostr$. It can be seen that if $X$ contains more than zero variables, then the function $f_{\psi(x,X)}$ grows exponentially. We have now that $\psi(x)$ contains only first order variables. Consider some $\A \in \ostr$ of size $n$ and let $a_1,\ldots,a_s\in \A$ be the tuples that satisfy $\A \models \psi(a_i)$. Now consider a structure $\A'$ that is obtained by duplicating $\A$. This is, $\A'$ has size $2n$, its first half is isomorphic to its second half, and also to $\A$. We have that there is at least $2s$ tuples $a_i,b_i\in \A'$, $s$ on each half, such that $\A'\models\psi(a_i)$ and $\A'\models\psi(b_i)$, which leads to a contradiction.

\textbf{Proof of Theorem \ref{mult-sigma-zero}}

Let $\alpha \in \sqso(\loge{0})$ with binary product. This is, $\alpha$ is defined by the grammar $\alpha := \sa{x}\varphi \ \mid \ \sa{X}\varphi \ \mid \ \sa{x}\alpha \ \mid \ \sa{X}\alpha \ \mid \ (\alpha\mult\alpha)$, where $\varphi \in \loge{0}$. For convenience, we add the expressions $\sa{x}(\varphi\mult\varphi)$ and $\sa{X}(\varphi\mult\varphi)$ to the grammar. Let $k$ be the number of algebraic operators in $\alpha$. We will show by induction on $k$ that $\alpha$ is expressible without multiplication. If $k = 1$, then the possible cases are $\alpha = \sa{x}\varphi$, where $\varphi\in\loge{0}$, which is already product-free. If $k = 2$, then the only that is not two nested summations is $\alpha = \sa{x}(\varphi\mult\psi)$, which can be expressed as $\alpha = \sa{x}(\varphi\wedge\psi)$. Suppose that the hypothesis holds for $k = 1$ through $n-1$. Let $\alpha$ have $n$ algebraic operators. If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$ where $\beta$ is an algebraic expression, then $\beta$ has $n-1$ algebraic operators and by hypothesis it can be expressed as product-free. If $\alpha = (\beta \mult \gamma)$ there are two cases. If at least one between $\beta$ and $\gamma$ has a product, then by hypothesis it can be expressed as product-free. Moreover, the new expression for $\alpha$ has strictly less than $n$ algebraic operators so by hypothesis it can be expressed as product-free. If there are no products in $\beta$ or $\gamma$, then let $\beta = \sa{x}\delta(x)$. We have that $\alpha$ can be expressed as $\sa{y}(\delta(y)\mult\gamma)$, where $y$ is a first-order variable that does not appear in $\beta$ or $\gamma$. Since $\delta(y)\mult\gamma$ has $n-1$ algebraic operators, it can be expressed as product-free, and then $\alpha$ can be expressed as product-free.

\textbf{Proof of Theorem \ref{fo-prod-sigma-zero}}

The resulting grammar is $\alpha := \sa{x}\varphi \ \mid \ \sa{X} \varphi \ \mid \ \sa{x}\alpha \ \mid \ \sa{X}\alpha \ \mid \pa{x}\alpha$. Consider a more restricted grammar that replaces $\pa{x}\alpha$ by $\pa{x}\sa{y}\varphi$. Let $\alpha = \pa{x}\sa{y}(x = y)$. This formula produces a constant function equal to 1, which as shown in Theorem \ref{one-sigma-zero} is not expressible in $\E{0}$.

\textbf{Proof of Theorem \ref{fp1}}

Let $f \in \XE{0}$, and let $\varphi(\X,\x)$ be and {\sc FO}-extended quantifier-free $\L$-formula such that:
\begin{eqnarray*}
	f(\A) &=& \vert \{\langle \P,\e  \rangle \mid \A \models \varphi(\P,\e) \} \vert
\end{eqnarray*}
for each $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\e \in A^m$ and $\P = (P_1,\ldots,P_q)$ is a tuple of predicates. We will now show that computing $f(\A)$ can be done in polynomial time.

For each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$, let $R_{\beta}$ be a predicate of arity $m$. Let $\A^\prime = \langle A, \S^{\A}, R_{\beta}^{\A^\prime}, \leq^{\A} \rangle \in \str$, where $R_{\beta}^{\A^\prime} = \{\d \mid \A \models \beta(\d)\}$. Note that each $\beta(\x)$ is fixed in $\varphi(\X,\x)$, and for each $\d \in A^m$, checking whether $\A \models \beta(\d)$ can be done in polynomial time. Therefore, generating $R_{\beta}^{\A^\prime}$ can also be done in polynomial time.

Let $\psi(\X,\x)$ be obtained by replacing each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$ by $R_{\beta}(\x)$. Also, let $g = f_{\psi(\X,\x)}$. Note that for each tuple of predicates $\P$ and each $\e \in A^m$, $\A \models \varphi(\P,\e)$ if and only if $\A^\prime \models \psi(\P,\e)$, and so, $g(\A^\prime) = f(\A)$. But, $\psi(\X,\x)$ is a quantifier-free $\L$-formula, and therefore, $g \in \#\Sigma_0$. Since it is shown in \cite{DBLP:journals/jcss/SalujaST95} that $\#\Sigma_0 \subseteq$ {\sc FP}, we conclude that $f(\A)$ can be evaluated in polynomial time.

@@

\textbf{Proof of Theorem \ref{sigmafo-minusone}}

Let $\alpha$ be a $\eqso(\logex{1})$ formula over a signature $\R$. We will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$. Let $\tau(\alpha) = \beta_1\add\cdots\add\beta_n$ be defined as in \ref{eqso-sigma-zero-in-fp}, that is, each $\beta_i$ is equal to $1$, $\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{x}}\sa{\bar{X}}1$ or $\sa{\bar{x}}\sa{\bar{X}}\varphi$, where $\varphi$ is a formula in $\logex{1}$ and $\bar{x}$ ($\bar{X}$) is a tuple of first-order (second-order) variables. We further assume that each $\beta_i$ is either $\sa{\bar{x}}\varphi$ or $\sa{\bar{x}}\sa{\bar{X}}\varphi$, since each $1$, $\varphi$ and $\sa{\bar{X}}\varphi$ can be replaced by the formulas $\sa{x}\min(x)$, $\sa{x}(\varphi\wedge\min(x))$ and $\sa{x}\sa{\bar{X}}(\varphi\wedge\min(x))$ respectively, while maintaining the value of $\sem{\alpha}(\A)$.

The intuition behind our reasoning is separated in three points: (1) For each $\beta_i$ of the form $\sa{\bar{x}}\varphi$, the formula $\kappa(\beta_i)$ will count every tuple $\bar{x}$ that satisfies $\varphi$ except for the lexicographically smallest one. (2) For each $\beta_i$ of the form $\sa{\bar{x}}\sa{\bar{X}}\varphi$, the formula $\kappa(\beta_i)$ will isolate the smallest $\bar{X}$ that satisfies $\varphi$, and exclude the lexicographically smallest tuple $\bar{x}$ that satisfies $\varphi(\bar{X})$. And (3) if $\alpha = (\beta\add\sa{\bar{x}}\varphi)$ or $\alpha = (\beta\add\sa{\bar{x}}\sa{\bar{X}}\varphi)$, the formula $\kappa(\alpha)$ will exclude the lexicographically smallest tuple that satisfies $\varphi$ if and only if $\beta$ is equal to 0.

@@

For each $\alpha$ in $\eqso(\logex{1})$ such that $\alpha = \sa{\bar{x}}\sa{\bar{X}}\exists\bar{y}\,\varphi(\bar{x},\bar{X},\bar{y})$ for some quantifier-free formula $\varphi$, we define $\kappa(\alpha)$ procedurally as follows: Let $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$ and $\bar{X} = (X_1,\ldots,X_{\length(\bar{X})})$. We suppose that $\varphi$ is in a DNF form that leaves formulas that do not mention $\bar{X}$ as is. If it is not, we convert $\varphi$ to this form with a standard transformation algorithm. Let $\varphi(\bar{x},\bar{X},\bar{y}) = \bigvee_{i = 1}^m\varphi_i(\bar{x},\bar{X},\bar{y})$ where each $\varphi_i$ has the form:
$$
\varphi_i(\bar{x},\bar{X},\bar{y}) = (\text{conjunction of $X_i$'s}) \wedge (\text{conjunction of $\neg X_i$'s})  \wedge (\text{an $\fo$ formula that does not mention any $X_i$}).
$$
Define $\varphi_i^{+}$, $\varphi_i^{-}$ and $\varphi_i^{\fo}$ as the formulas mentioned above.

In our procedure we assume that in $\varphi_i^{+}\wedge\varphi_i^{-}$ every first-order variable from $\bar{x},\bar{y}$ is mentioned at most once. If not, we add new first-order variables from $\fv$ to $\bar{y}$ so that no variable is repeated, replace them in $\varphi_i^{+}\wedge\varphi_i^{-}$ and add their respective equalities in $\varphi_i^{\fo}$ so that the new formula is equivalent. For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (x_1,x_2,x_3,x_4,y)$ and $\varphi_i := X(x_1,x_2) \wedge \neg X(x_3,x_4) \wedge x < y \wedge x_1 = x \wedge x_2 = y \wedge x_3 = x \wedge x_4 = x.$

>>>>

First, for a given signature $\R$ and logic fragment $\LL$ we define a {\em valid form in $\LL$} (or just {\em valid form}). A formula $\alpha$ is in valid form in $\LL$ if (1) it is equal to $\varphi$, $1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}1$ for some $\LL$ formula $\varphi$ and some tuple of first-order (second-order) variables $\bar{x}$ ($\bar{X}$), or (2) if it is equal to $(\beta\add\gamma)$ where $\beta$ and $\gamma$ are formulas in valid form. Given a $\eqso(\loge{0})$ formula $\alpha$ over a signature $\R$ we will recursively construct a formula $\tau(\alpha)$ in valid form such that $\alpha(\enc(\A)) = \tau(\alpha)(\enc(\A))$ for each $\A\in\ostr[\R]$.

First, each natural constant $s$ in $\alpha$ is replaced by $(1\add 1\add\cdots\add 1)$ ($s$ times). If $\alpha = 1$ or $\alpha = \varphi \in \loge{0}$, then $\tau(\alpha) = \alpha$. We assume that every sub-formula of $\alpha$ is in valid form. If $\alpha = (\beta \add \gamma)$ then $\tau(\alpha) = \alpha$. If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$ this case separates in more cases: If $\alpha = \sa{x}\beta$ and $\beta$ is in one of the valid forms in (1), then $\tau(\alpha) = \alpha$. If $\alpha = \sa{X}\beta$ and $\beta$ is in the valid form $\sa{\bar{x}}\gamma$ then $\tau(\alpha) = \sa{\bar{x}}\sa{X}\gamma$; if $\beta$ it is in one of the remaining valid forms in (1), then $\tau(\alpha) = \alpha$. If $\alpha = \sa{x}\beta$ ($\alpha = \sa{X}\beta$) and $\beta = (\gamma_1\add\gamma_2)$, then $\tau(\alpha) = \tau(\sa{x}\gamma_1) \add \tau(\sa{x}\gamma_2)$ ($\tau(\alpha) = \tau(\sa{X}\gamma_1) \add \tau(\sa{X}\gamma_2)$). This covers all possible cases for $\alpha$. Furthermore, suppose that $\LL$ is closed under conjunction and suppose $\alpha = (\beta\mult\gamma)$. If $\beta = (\delta_1\add\delta_2)$ (wlog), then $\tau(\alpha) = \tau(\delta_1\mult\gamma)\add\tau(\delta_2\mult\gamma)$. If $\beta = \sa{\bar{x}}\delta$, then $\tau(\alpha) = \tau(\sa{\bar{x}}\tau(\delta\mult\gamma))$. If $\beta = 1$, then $\tau(\alpha) = \gamma$. If $\beta = \varphi$ and $\gamma = \psi$, then $\tau(\alpha) = \varphi\wedge\psi$. Let $\alpha$ be a $\eqso(\loge{0})$ formula over a signature $\R$. Consider the following procedure that receives $\enc(\A)$ as input, where $\A\in\ostr[\R]$ with domain $A$. Let $\tau(\alpha) = \beta_1\add\cdots\add\beta_m$. For each $\beta_i$, if $\beta_i = 1$ or $\beta_i = \varphi$, then $\sem{\beta_i}(\A)$ can be computed in polynomial time. If $\beta_i = \sa{\bar{x}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}$, if $\beta_i = \sa{\bar{X}}1$ then $\sem{\beta_i}(\A) = 2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, if $\beta_i = \sa{\bar{X}}\sa{\bar{x}}1$ then $\sem{\beta_i}(\A) = \vert A \vert^{\length(\bar{x})}2^{\Sigma_{X\in\bar{X}}\vert A \vert^{\arity(X)}}$, all of which can be computed in polynomial time. Otherwise, $\beta_i$ is $\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, and then $\sem{\beta_i}(\A) = f_{\varphi}(\enc(\A))$, and $f_{\varphi}\in \#\loge{0}$ which implies that $f_{\varphi}(\enc(\A))$ can be computed in polynomial time \cite{DBLP:journals/jcss/SalujaST95}.

@@

Let $\alpha$ be a formula in $\eqso(\loge{1})$ over a signature $\R$. Let $\tau(\alpha) = \beta_1\add\beta_2\add\cdots\add\beta_m$ defined as in the proof of Theorem \ref{eqso-sigma-zero-in-fp}. This is, each $\beta_i$ is of the form $1$, $\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{X}}\sa{\bar{x}}1$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, where $\varphi$ is a formula in $\loge{1}$ and $\bar{x}$ ($\bar{X}$) is a tuple of first-order (second-order) variables. For each $\beta_i$ we will construct a $\gamma_i$ such that $\gamma_i \in \eqso(\logu{1})$ and $\sem{\beta_i}(\A) = \sem{\gamma_i}(\A)$ for each $\A\in\ostr[\R]$. Since the forms $1$, $\sa{\bar{x}}1$, $\sa{\bar{X}}1$ and $\sa{\bar{X}}\sa{\bar{x}}1$ are already valid in $\eqso(\logu{1})$, we suppose $\beta_i$ is in one of the remaining forms. Without loss of generality, let $\beta_i = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{x},\bar{y},\bar{X})$ where $\varphi(\bar{x},\bar{y},\bar{X})$ is a quantifier-free formula. Similarly as in the proof by Saluja et al. \cite{DBLP:journals/jcss/SalujaST95}, we define $\gamma_i = \sa{\bar{x}}\sa{\bar{y}}\sa{\bar{X}}\varphi(\bar{x},\bar{y},\bar{X}) \wedge \forall\bar{z}(\varphi(\bar{x},\bar{z},\bar{X})\to(\bar{z} < \bar{y} \vee \bar{z} = \bar{y}))$ that fulfils the desired condition. The theorem holds true by using $\beta = \gamma_1\add\cdots\add\gamma_m$ which is in $\eqso(\logu{1})$

%\subsection{Proof of Theorem \ref{one-sigma-zero}}
%
%Towards a contradiction, suppose that there exists a constant function $f(\A) = s > 0$ which is expressible in $\E{0}$. This is, there is a formula $\psi(x,X)$ such that $f_{\psi(x,X)}(\A) = \vert \{ \langle e,P\rangle \mid \A\models\psi(e,P) \} \vert = s$ for each $\A\in\ostr$. It can be seen that if $X$ contains more than zero variables, then the function $f_{\psi(x,X)}$ grows exponentially. We have now that $\psi(x)$ contains only first order variables. Consider some $\A \in \ostr$ of size $n$ and let $a_1,\ldots,a_s\in \A$ be the tuples that satisfy $\A \models \psi(a_i)$. Now consider a structure $\A'$ that is obtained by duplicating $\A$. This is, $\A'$ has size $2n$, its first half is isomorphic to its second half, and also to $\A$. We have that there is at least $2s$ tuples $a_i,b_i\in \A'$, $s$ on each half, such that $\A'\models\psi(a_i)$ and $\A'\models\psi(b_i)$, which leads to a contradiction.

%\textbf{Proof of Theorem \ref{mult-sigma-zero}}
%
%Let $\alpha \in \sqso(\loge{0})$ with binary product. This is, $\alpha$ is defined by the grammar $\alpha := \sa{x}\varphi \ \mid \ \sa{X}\varphi \ \mid \ \sa{x}\alpha \ \mid \ \sa{X}\alpha \ \mid \ (\alpha\mult\alpha)$, where $\varphi \in \loge{0}$. For convenience, we add the expressions $\sa{x}(\varphi\mult\varphi)$ and $\sa{X}(\varphi\mult\varphi)$ to the grammar. Let $k$ be the number of algebraic operators in $\alpha$. We will show by induction on $k$ that $\alpha$ is expressible without multiplication. If $k = 1$, then the possible cases are $\alpha = \sa{x}\varphi$, where $\varphi\in\loge{0}$, which is already product-free. If $k = 2$, then the only that is not two nested summations is $\alpha = \sa{x}(\varphi\mult\psi)$, which can be expressed as $\alpha = \sa{x}(\varphi\wedge\psi)$. Suppose that the hypothesis holds for $k = 1$ through $n-1$. Let $\alpha$ have $n$ algebraic operators. If $\alpha = \sa{x}\beta$ or $\alpha = \sa{X}\beta$ where $\beta$ is an algebraic expression, then $\beta$ has $n-1$ algebraic operators and by hypothesis it can be expressed as product-free. If $\alpha = (\beta \mult \gamma)$ there are two cases. If at least one between $\beta$ and $\gamma$ has a product, then by hypothesis it can be expressed as product-free. Moreover, the new expression for $\alpha$ has strictly less than $n$ algebraic operators so by hypothesis it can be expressed as product-free. If there are no products in $\beta$ or $\gamma$, then let $\beta = \sa{x}\delta(x)$. We have that $\alpha$ can be expressed as $\sa{y}(\delta(y)\mult\gamma)$, where $y$ is a first-order variable that does not appear in $\beta$ or $\gamma$. Since $\delta(y)\mult\gamma$ has $n-1$ algebraic operators, it can be expressed as product-free, and then $\alpha$ can be expressed as product-free.
%
%\textbf{Proof of Theorem \ref{fo-prod-sigma-zero}}
%
%The resulting grammar is $\alpha := \sa{x}\varphi \ \mid \ \sa{X} \varphi \ \mid \ \sa{x}\alpha \ \mid \ \sa{X}\alpha \ \mid \pa{x}\alpha$. Consider a more restricted grammar that replaces $\pa{x}\alpha$ by $\pa{x}\sa{y}\varphi$. Let $\alpha = \pa{x}\sa{y}(x = y)$. This formula produces a constant function equal to 1, which as shown in Theorem \ref{one-sigma-zero} is not expressible in $\E{0}$. 
%
%\textbf{Proof of Theorem \ref{fp1}}
%
%Let $f \in \XE{0}$, and let $\varphi(\X,\x)$ be and {\sc FO}-extended quantifier-free $\L$-formula such that:
%\begin{eqnarray*}
%	f(\A) &=& \vert \{\langle \P,\e  \rangle \mid \A \models \varphi(\P,\e) \} \vert
%\end{eqnarray*}
%for each $\A = \langle A, \S^{\A}, \leq^{\A} \rangle \in \str$, where $\e \in A^m$ and $\P = (P_1,\ldots,P_q)$ is a tuple of predicates. We will now show that computing $f(\A)$ can be done in polynomial time.
%
%For each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$, let $R_{\beta}$ be a predicate of arity $m$. Let $\A^\prime = \langle A, \S^{\A}, R_{\beta}^{\A^\prime}, \leq^{\A} \rangle \in \str$, where $R_{\beta}^{\A^\prime} = \{\d \mid \A \models \beta(\d)\}$. Note that each $\beta(\x)$ is fixed in $\varphi(\X,\x)$, and for each $\d \in A^m$, checking whether $\A \models \beta(\d)$ can be done in polynomial time. Therefore, generating $R_{\beta}^{\A^\prime}$ can also be done in polynomial time.
%
%Let $\psi(\X,\x)$ be obtained by replacing each {\sc FO}-formula $\beta(\x)$ in $\varphi(\X,\x)$ by $R_{\beta}(\x)$. Also, let $g = f_{\psi(\X,\x)}$. Note that for each tuple of predicates $\P$ and each $\e \in A^m$, $\A \models \varphi(\P,\e)$ if and only if $\A^\prime \models \psi(\P,\e)$, and so, $g(\A^\prime) = f(\A)$. But, $\psi(\X,\x)$ is a quantifier-free $\L$-formula, and therefore, $g \in \#\Sigma_0$. Since it is shown in \cite{DBLP:journals/jcss/SalujaST95} that $\#\Sigma_0 \subseteq$ {\sc FP}, we conclude that $f(\A)$ can be evaluated in polynomial time.

@@

\subsection{Proof of Theorem \ref{sigmafo-minusone}}

Let $\alpha$ be a $\eqso(\logex{1})$ formula over a signature $\R$. We will define a $\eqso(\logex{1})$ formula $\kappa(\alpha)$ such that for each finite structure $A$ over $\R$: $\sem{\kappa(\alpha)}(\A) = \sem{\alpha}(\A) \dotminus 1$. Let $\tau(\alpha) = \beta_1\add\cdots\add\beta_n$ be defined as in \ref{eqso-sigma-zero-in-fp}, that is, each $\beta_i$ is equal to $1$, $\varphi$, $\sa{\bar{x}}1$, $\sa{\bar{x}}\varphi$, $\sa{\bar{X}}1$, $\sa{\bar{X}}\varphi$, $\sa{\bar{X}}\sa{\bar{x}}1$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, where $\varphi$ is a formula in $\logex{1}$ and $\bar{x}$ ($\bar{X}$) is a tuple of first-order (second-order) variables. We further assume that each $\beta_i$ is either $\sa{\bar{x}}\varphi$ or $\sa{\bar{X}}\sa{\bar{x}}\varphi$, since each $1$, $\varphi$ and $\sa{\bar{X}}\varphi$ can be replaced by the formulas $\sa{x}\min(x)$, $\sa{x}(\varphi\wedge\min(x))$ and $\sa{x}\sa{\bar{X}}(\varphi\wedge\min(x))$ respectively, while maintaining the value of $\sem{\alpha}(\A)$.

The intuition behind our reasoning is separated in three points: (1) For each $\beta_i$ of the form $\sa{\bar{x}}\varphi$, the formula $\kappa(\beta_i)$ will count every tuple $\bar{x}$ that satisfies $\varphi$ except for the lexicographically smallest one. (2) For each $\beta_i$ of the form $\sa{\bar{X}}\sa{\bar{x}}\varphi$, the formula $\kappa(\beta_i)$ will isolate the smallest $\bar{X}$ that satisfies $\varphi$, and exclude the lexicographically smallest tuple $\bar{x}$ that satisfies $\varphi(\bar{X})$. And (3) if $\alpha = (\beta\add\sa{\bar{x}}\varphi)$ or $\alpha = (\beta\add\sa{\bar{X}}\sa{\bar{x}}\varphi)$, the formula $\kappa(\alpha)$ will exclude the lexicographically smallest tuple that satisfies $\varphi$ if and only if $\beta$ is equal to 0.

\vspace{1em}

@@

\vspace{1em}

For each $\alpha$ in $\eqso(\logex{1})$ such that $\alpha = \sa{\bar{X}}\sa{\bar{x}}\exists\bar{y}\,\varphi(\bar{X},\bar{x},\bar{y})$ for some quantifier-free formula $\varphi$, we define $\kappa(\alpha)$ procedurally as follows: Let $\bar{x} = (x_1,\ldots,x_{\length(\bar{x})})$ and $\bar{X} = (X_1,\ldots,X_{\length(\bar{X})})$. We suppose that $\varphi$ is in a DNF form that leaves formulas that do not mention $\bar{X}$ as is. If it is not, we convert $\varphi$ to this form with a standard transformation algorithm. Let $\varphi(\bar{X},\bar{x},\bar{y}) = \bigvee_{i = 1}^m\varphi_i(\bar{X},\bar{x},\bar{y})$ where each $\varphi_i$ has the form:
$$
\varphi_i(\bar{X},\bar{x},\bar{y}) = (\text{conjunction of $X_i$'s}) \wedge (\text{conjunction of $\neg X_i$'s})  \wedge (\text{an $\fo$ formula that does not mention any $X_i$}).
$$
Define $\varphi_i^{+}$, $\varphi_i^{-}$ and $\varphi_i^{\fo}$ as the formulas mentioned above.

In our procedure we assume that in $\varphi_i^{+}\wedge\varphi_i^{-}$ every first-order variable from $\bar{x},\bar{y}$ is mentioned at most once. If not, we add new first-order variables from $\fv$ to $\bar{y}$ so that no variable is repeated, replace them in $\varphi_i^{+}\wedge\varphi_i^{-}$ and add their respective equalities in $\varphi_i^{\fo}$ so that the new formula is equivalent. For example, if $\bar{x} = x$, $\bar{y} = y$ and $\varphi_i = X(x,y)\wedge \neg X(x,x) \wedge x < y$, then we redefine $\bar{y} = (y,v_1,v_2,v_3,v_4)$ and $\varphi_i := X(v_1,v_2) \wedge \neg X(v_3,v_4) \wedge x < y \wedge v_1 = x \wedge v_2 = y \wedge v_3 = x \wedge v_4 = x.$

Furthermore, we assume that $\varphi_i^{\fo}$ defines an ordered partition of the variables in the tuple $(\bar{x},\bar{y})$. For example, let $\bar{x} = (x_1,x_2,x_3,x_4)$. A possible ordered partition would be defined by the formula $\theta(\bar{x}) = x_2 < x_1 \wedge x_1 = x_4 \wedge x_4 < x_3$. On the other hand, the formula $\theta'(\bar{x}) = x_1 < x_2 \wedge x_1 < x_4 \wedge x_2 = x_3$ does not define an ordered partition since both $\{x_1\}<\{x_2,x_3\}<\{x_4\}$ and $\{x_1\} < \{x_2,x_3,x_4\}$ satisfy $\theta'$.
For a given $k$, we define $\cB_k$ as the number of possible ordered partitions for a set of size $k$. 
If $\varphi_i^{\fo}$ does not define an ordered partition, then for $1 \leq j \leq \cB_{\length(\bar{x},\bar{y})}$ 
let $\theta^j(\bar{x},\bar{y})$ be a formula that defines an ordered partition over $(\bar{x},\bar{y})$. The formula $\varphi_i(\bar{X},\bar{x},\bar{y})$ is converted into $\bigvee_{j = 1}^{\cB_{\length(\bar{x},\bar{y})}}\varphi_i(\bar{X},\bar{x},\bar{y}) \wedge \theta^j(\bar{x},\bar{y})$. Consider each of these disjuncts as a new $\varphi_i$.

After the previous assumptions, we do the following: for each $X_j\in \bar{X}$ we check every instance of $X_j(\bar{w})$ in $\varphi^{+}_i$ and every instance of $X_j(\bar{z})$ in $\varphi^{-}_i$, where $\bar{w}$ and $\bar{z}$ are subtuples of $(\bar{x},\bar{y})$. If the ordered partition in $\varphi^{\fo}_i$ satisfies $\bar{w} = \bar{z}$, the entire formula $\varphi_i$ is removed from $\varphi$.
